Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"

### 2020-02-24 writing/anderson_ehp_supp.tex
Work on some final edits throughout

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{othertopstorms}
\caption{Differences in the counties assessed as ``exposed'', based on different
	exposure metrics, for a sample of storms. These sample storms were 
	selected as the  storms with largest extent (as measured by the 
	number of counties exposed based on any metric) from each of the 
	clusters shown in the Jaccard heatmap in the main text (Figure 7).}
\label{fig:othertopstorms}
\end{figure}

>>>>

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{othertopstorms}
\caption{Differences in the counties assessed as ``exposed'', based on different
	exposure metrics, for a sample of storms. These sample storms were 
	selected as the  storms with largest extent (as measured by the 
	number of counties exposed based on any metric) from each of the 
	clusters shown in the Jaccard heatmap in the main text (Figure 7; a similar
	map for Hurricane Ivan in 2004 is shown in Figure 6 of the main text).}
\label{fig:othertopstorms}
\end{figure}

### 2020-02-24 writing/ehp_figures.tex
Work on some final edits throughout

\begin{figure}[tbhp!]
\centering
\includegraphics[width=0.8\linewidth]{windcomparison}
	\caption{Comparison of two sources of wind exposure estimates: 
	(1) modeled peak sustained surface wind 
	and (2) estimates based on \ac{HURDAT2}'s wind speed radii.
	Each point represents a storm, with the x axis giving the percent of
	counties classified in the same category of peak sustained surface
	wind ($<$34~kt; 34\,--\,49.9~kt;
	50\,--\,63.9~kt; $\ge$64~kt) by both sources of 
	data. The color of each point gives the number of study counties that were exposed to
	peak sustained surface wind of at least 34~kt (based
	on modeled wind). Estimates are shown for study storms since
	2004, the earliest year for which post-storm reanalysis wind speed
	radii estimates are routinely available in \ac{HURDAT2}, and for which at
	least one study county had a peak sustained wind of $\ge$34~\si{\knot}
	based on the \ac{HURDAT2} wind radii.
	}
\label{fig:windcomparison}
\end{figure}

@@

\begin{figure}%[tbhp] 
\centering
\includegraphics[width=7cm]{averageexposureonly.pdf} 
\caption{Average number of county-level storm exposures per decade for binary 
	classifications based on each
	single-hazard exposure metric (Table \ref{tab:exposuremetrics}). The years used to
	estimate these averages are based on years of available exposure data
	(rain: 1988\,--\,2011; wind:~1988\,--\,2018; flood events:~1996\,--\,2018, and
	tornado events:~1988\,--\,2018). } 
\label{fig:averageexposure} 
\end{figure}

@@

\begin{figure}%[tbhp] 
	\centering 
	\includegraphics[width = 0.8\linewidth]{jaccard_heatmap.pdf}
	\caption{Agreement between exposure
	classifications based on different exposure metrics for all storms
	between~1996 and~2011 for which at least~100 counties were exposed
	based on at least one metric.  Each row shows one storm, and the color
	of each cell shows the measured Jaccard index for each pair of exposure
	metrics (proportion of counties classified as exposed by both metrics
	out of counties classified as exposed by either metric). For 
	Grace in 2003 and Ida in 2009, there were no county exposures for
	either the tornado-based metric or the wind-based metric (indicated 
	by gray squares). Colors to the right of the
	heatmap show the number of exposed counties
	based on any of the metrics. Storms are displayed
	within clusters that have similar patterns based on hierarchical 
	clustering.}
\label{fig:jaccard} 
\end{figure}

>>>>

\begin{figure}[tbhp!]
\centering
\includegraphics[width=0.8\linewidth]{windcomparison}
	\caption{Comparison of two sources of wind exposure estimates: 
	(1) modeled peak sustained surface wind 
	and (2) estimates based on \ac{HURDAT2}'s wind speed radii.
	Each point represents a storm, with the x-axis giving the percent of
	counties classified in the same category of peak sustained surface
	wind ($<$34~kt; 34\,--\,49.9~kt;
	50\,--\,63.9~kt; $\ge$64~kt) by both sources of 
	data. The color of each point gives the number of study counties that were exposed to
	peak sustained surface wind of at least 34~kt (based
	on modeled wind). Estimates are shown for study storms since
	2004, the earliest year for which post-storm reanalysis wind speed
	radii are routinely available in \ac{HURDAT2}, and for which at
	least one study county had a peak sustained wind of $\ge$34~\si{\knot}
	based on the post-storm wind radii.
	}
\label{fig:windcomparison}
\end{figure}

@@

\begin{figure}%[tbhp] 
\centering
\includegraphics[width=7cm]{averageexposureonly.pdf} 
\caption{Average number of county-level storm exposures per decade for binary 
	classifications based on each
	single-hazard exposure metric (Table \ref{tab:exposuremetrics}). The years used to
	estimate these averages are based on years of available exposure data
	(rain: 1988\,--\,2011; wind:~1988\,--\,2018; flood events:~1996\,--\,2018; and
	tornado events:~1988\,--\,2018). } 
\label{fig:averageexposure} 
\end{figure}

@@

\begin{figure}%[tbhp] 
	\centering 
	\includegraphics[width = 0.8\linewidth]{jaccard_heatmap.pdf}
	\caption{Agreement between exposure
	classifications based on different single-hazard exposure metrics for all storms
	between~1996 and~2011 for which at least~100 counties were exposed
	based on at least one metric.  Each row shows one storm, and the color
	of each cell shows the measured Jaccard index for each pair of exposure
	metrics (proportion of counties classified as exposed by both metrics
	out of counties classified as exposed by either metric). For 
	Grace in 2003 and Ida in 2009, there were no county exposures for
	either the tornado-based metric or the wind-based metric (indicated 
	by gray squares). Colors to the right of the
	heatmap show the number of exposed counties
	based on any of the metrics. Storms are displayed
	within clusters that have similar patterns based on hierarchical 
	clustering.}
\label{fig:jaccard} 
\end{figure}

### 2020-02-24 writing/ehp_introduction.tex
Work on some final edits throughout

Hurricanes and other tropical cyclones can severely impact human health in
\ac{US} communities, as shown in several long-term national
studies~\parencite{rappaport2000, rappaport2014fatalities,
rappaport2016fatalities, czajkowski2010fatal, czajkowski2011, moore2012}.
These studies characterize tropical cyclone health impacts using fatality data
aggregated in large part from disaster-related mortality ascertainment and
surveillance conducted by agencies like the \ac{US} Centers for Disease Control
and Prevention, the \ac{US} National Weather Service, and local vital
statistics departments. These fatality data are based on case-by-case
ascertainment: identifying and characterizing deaths for which there is a clear
link with a disaster, through death certificate coding or other indicators
\parencite{rocha2017medicolegal}.

@@

The \ac{NHC} publishes a ``Best Tracks'' dataset that is considered the gold
standard for Atlantic-basin tropical cyclone tracks.  It records the central
position of a tropical cyclone every six hours, as well as the storm's minimum
central pressure and maximum sustained surface wind.  These data are openly
available through the \ac{HURDAT2}, a post-storm assessment that incorporates
data from several sources, including satellite data and, when available,
aircraft reconnaissance data~\parencite{landsea2013, jarvinen1988}.  With these
data, it is straightforward to measure a storm's direct path, and so to measure
whether a community was on or within a certain distance of that central track.
Some previous epidemiological studies have done this to assess exposure to a
tropical storm (e.g.,~\cite{currie2013, kinney2008, caillouet2008increase}), 
and this method can be applied consistently across years and communities 
for large-scale studies.

However, when epidemiological studies assess exposure to a tropical cyclone
using, as a proxy, how close the storm came to a community, they may
misclassify exposure~\parencite{grabich2015measuring}. While a number of
tropical cyclone hazards are strongly associated with distance from the
tropical cyclone's center (e.g., wind and, at the coast, storm surge and
waves~\parencite{rappaport2000, kruk2010}), other hazards like heavy rainfall,
floods, and tornadoes can occur well away from the tropical cyclone's central
track~\parencite{rappaport2000, atallah2007, moore2012}.  For example,
tornadoes generated by tropical cyclones, which were linked to over~300 deaths
in the \ac{US} between~1995 and~2009, most often
occur~200\,--\,500~\si{\kilo\metre} from the tropical cyclone's
center~\parencite{moore2012}.  Further, when studies use distance from the
storm's track to assess exposure, they often use the same buffer constraints on
each side of the storm track (e.g.,~\cite{kinney2008, currie2013}).  However,
the forces of a tropical cyclone tend to be distributed around its center
non-symmetrically.  Extreme winds are more common to the track's right, where
counter-clockwise cyclonic winds move in concert with the tropical cyclone's
forward motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
associated with \ac{US} tropical cyclones between~1995 and~2009 occurred to the
right of the tropical cyclone's track~\parencite{moore2012}. Rain, conversely,
is often heaviest to the left of the track, especially when the tropical
cyclone interacts with other weather systems~\parencite{atallah2003,
atallah2007, zhu2013variations} or undergoes an extratropical
transition~\parencite{elsberry2002}.  The multi-hazard nature of tropical
cyclones therefore makes it hard to assess exposure based on how close the
storm's central track came to the community.  While other approaches have been
developed to incorporate storm hazards, particularly wind, into exposure
assessment (e.g., \cite{grabich2015measuring, zandbergen2009, czajkowski2011}),
there is not yet a standard approach, and when different studies use
different datasets or storm hazards when assessing storm exposure, it becomes
difficult to compare and aggregate findings.

To help epidemiological researchers assess exposure to tropical cyclones, we
developed an open-source dataset~\parencite{hurricaneexposure}, which we
present here.  These data cover all counties in states in the eastern half of
the \ac{US} over multiple years and include five metrics characterizing
exposure to tropical cyclones (Table~\ref{tab:exposuremetrics}): (1)~closest
distance the storm's central track came to the county's center (a proxy
measurement of storm exposure used in some previous studies); (2)~peak
sustained surface wind at the county's center over the course of the storm;
(3)~cumulative rainfall in the county over the course of the storm;
(4)~flooding in the county concurrent with the storm; and (5)~tornadoes in the
county concurrent with the storm.  We aggregated these data at the county level
since data for epidemiological studies are often available at this level (e.g.,
direct hurricane-related deaths~\parencite{czajkowski2011}; birth
outcomes~\parencite{grabich2015, grabich2015measuring}; autism
prevalence~\parencite{kinney2008}) and since decisions and policies to prepare
for, and respond to, tropical cyclones are often undertaken at the county
level~\parencite{zandbergen2009, rappaport2000}.  We also wrote functions to
explore and map these data and to link them with county-level health
data~\parencite{hurricaneexposuredata}.

Here, we describe how we developed this dataset and explore how its
measurements compare to other data that could be used to characterize tropical
cyclone hazards for epidemiologic research. Further, we expand on previous
research on methods to measure exposure to tropical cyclones for epidemiologic
research~\parencite{grabich2015measuring}. This previous study investigated
differences in which communities were assessed as exposed to tropical cyclones
during the 2004 hurricane season in Florida, comparing exposure assessment
based on distance to the storm's track or versus assessment based on a metric
that incorporated storm-generated winds within the county.  They found
important differences across methods of assessing storm exposure, concluding
that a study may be prone to bias from exposure misclassification if distance
to the storm track is used as a proxy measurement for exposure
assessment~\parencite{grabich2015measuring}.  Here we expand to investigate
this question across a larger set of counties and storm seasons.  Further, we
investigate patterns in exposure classification based on other metrics of
storm-related hazards---rainfall, flooding, and tornadoes---all of which are
important for inland health impacts of tropical
cyclones~\parencite{czajkowski2011, moore2012}. These results can help
epidemiologists design studies and plan statistical analysis for multi-year,
multi-community studies that estimate excess mortality and morbidity associated
with exposure to the hazards brought by tropical cyclones.

>>>>

Hurricanes and other tropical cyclones can severely impact human health in
\ac{US} communities, as shown in several long-term national
studies~\parencite{rappaport2000, rappaport2014fatalities,
rappaport2016fatalities, czajkowski2010fatal, czajkowski2011, moore2012}.
These studies characterize tropical cyclone health impacts using fatality data
aggregated in large part from disaster-related mortality ascertainment and
surveillance conducted by agencies like the \ac{US} Centers for Disease Control
and Prevention, the \ac{US} National Weather Service, and local vital
statistics departments. These fatality data are based on case-by-case
ascertainment: identifying and characterizing deaths for which there is a clear
link with a disaster through death certificate coding or other indicators
\parencite{rocha2017medicolegal}.

@@

The \ac{NHC} publishes a ``Best Tracks'' dataset that is considered the gold
standard for Atlantic-basin tropical cyclone tracks.  It records the central
position of a tropical cyclone every six hours, as well as the storm's minimum
central pressure and maximum sustained surface wind.  These data are openly
available through the \ac{HURDAT2}, a post-storm assessment that incorporates
data from several sources, including satellite data and, when available,
aircraft reconnaissance data~\parencite{landsea2013, jarvinen1988}.  With these
data, it is straightforward to measure a storm's direct path, and so to measure
whether a community was on or within a certain distance of that central track.
Some previous epidemiologic studies have done this to assess exposure to a
tropical storm (e.g.,~\cite{currie2013, kinney2008, caillouet2008increase}), 
and this method can be applied consistently across years and communities 
for large-scale studies.

However, when epidemiologic studies assess exposure to a tropical cyclone
using, as a proxy, how close the storm came to a community, they may
misclassify exposure~\parencite{grabich2015measuring}. While a number of
tropical cyclone hazards are strongly associated with distance from the
tropical cyclone's center (e.g., wind and, at the coast, storm surge and
waves~\parencite{rappaport2000, kruk2010}), other hazards like heavy rainfall,
floods, and tornadoes can occur well away from the tropical cyclone's central
track~\parencite{rappaport2000, atallah2007, moore2012}.  For example,
tornadoes generated by tropical cyclones, which were linked to over~300 deaths
in the \ac{US} between~1995 and~2009, most often
occur~200\,--\,500~\si{\kilo\metre} from the tropical cyclone's
center~\parencite{moore2012}.

Further, when studies use distance from the storm's track to assess exposure,
they often use the same buffer constraints on each side of the storm track
(e.g.,~\cite{kinney2008, currie2013}).  However, the forces of a tropical
cyclone tend to be distributed around its center non-symmetrically.  Extreme
winds are more common to the track's right, where counterclockwise cyclonic
winds move in concert with the tropical cyclone's forward
motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
associated with \ac{US} tropical cyclones between~1995 and~2009 occurred to the
right of the tropical cyclone's track~\parencite{moore2012}. Rain, conversely,
is often heaviest to the left of the track, especially when the tropical
cyclone interacts with other weather systems~\parencite{atallah2003,
atallah2007, zhu2013variations} or undergoes an extratropical
transition~\parencite{elsberry2002}.

The multi-hazard nature of tropical cyclones therefore makes it hard to assess
exposure based on how close the storm's central track came to the community.
While other approaches have been developed to incorporate storm hazards,
particularly wind, into exposure assessment (e.g., \cite{grabich2015measuring,
zandbergen2009, czajkowski2011}), there is not yet a standard approach, and
when different studies use different datasets or storm hazards when assessing
storm exposure, it becomes difficult to compare and aggregate findings.

To help epidemiological researchers assess exposure to tropical cyclones, we
developed an open-source dataset~\parencite{hurricaneexposure}, which we
present here.  These data cover all counties in states in the eastern half of
the \ac{US} over multiple years and include five metrics characterizing
exposure to tropical cyclones (Table~\ref{tab:exposuremetrics}): (1)~closest
distance the storm's central track came to the county's center (a proxy
measurement of storm exposure used in some previous studies); (2)~peak
sustained surface wind at the county's center over the course of the storm;
(3)~cumulative rainfall in the county over the course of the storm;
(4)~flooding in the county concurrent with the storm; and (5)~tornadoes in the
county concurrent with the storm.  We aggregated these data at the county level
since data for epidemiologic studies are often available at this level (e.g.,
direct hurricane-related deaths~\parencite{czajkowski2011}; birth
outcomes~\parencite{grabich2015, grabich2015measuring}; autism
prevalence~\parencite{kinney2008}) and since decisions and policies to prepare
for, and respond to, tropical cyclones are often undertaken at the county
level~\parencite{zandbergen2009, rappaport2000}.  We also wrote functions to
explore and map these data and to link them with county-level health
data~\parencite{hurricaneexposuredata}.

Here, we describe how we developed this dataset and explore how its
measurements compare to other data that could be used to characterize tropical
cyclone hazards for epidemiologic research. Further, we expand on previous
research on measuring exposure to tropical cyclones for epidemiologic
research~\parencite{grabich2015measuring}. This previous study investigated
differences in which communities were assessed as exposed to tropical cyclones
during the 2004 hurricane season in Florida, comparing exposure assessment
based on distance to the storm's track or versus a metric that incorporated
storm-generated winds within the county.  They found important differences,
concluding that a study may be prone to bias from exposure misclassification if
distance to the storm track is used as a proxy for
exposure~\parencite{grabich2015measuring}.  Here we expand to investigate this
question across a larger set of counties and storm seasons.  Further, we
investigate patterns in exposure classification based on other metrics of
storm-related hazards---rainfall, flooding, and tornadoes---important for
inland health impacts of tropical cyclones~\parencite{czajkowski2011,
moore2012}. These results can help epidemiologists design studies and plan
statistical analysis for multi-year, multi-community studies that estimate
excess mortality and morbidity associated with exposure to the hazards brought
by tropical cyclones.

### 2020-02-24 writing/ehp_methods.tex
Work on some final edits throughout

For our study domain, we aggregated data for all counties in the eastern half
of the \ac{US} (Figure~\ref{fig:hurrtracks}).  In terms of storm coverage, we
included all tropical cyclones between 1988 and 2018 in \ac{HURDAT2}
\parencite{landsea2013} that came within~250 \si{\kilo\metre} of at least one
eastern \ac{US} county, therefore including all land-falling or near-land
Atlantic basin tropical cyclones while excluding storms that never neared the
\ac{US}. The tracks of the study storms are shown in
Figure~\ref{fig:hurrtracks}).

\subsection*{Distance-based exposure metric}

We first measured how close each storm's central track came to each county. We
used tracking data from \ac{HURDAT2}, which records the storm center's position
at the synoptic times of 6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}.  We
interpolated this position to~15-\si{\minute} intervals using natural cubic
spline interpolation~\parencite{hurricaneexposure}. At each~15-\si{\minute}
interval, we measured the distance between the storm's center and each county's
population mean center~\parencite{bivand2013applied, countycenters}. We took
the minimum distance for each county as a measure of how close the storm came
to the county over its lifetime.  We also recorded the time when this occurred
for each county, which we later used to link observed data on precipitation,
flooding, and tornadoes.  To allow matching with data recorded in
local time (e.g., health data), we converted these times from \ac{UTC} to local
time~\parencite{countytimezones}.

@@

We based storm rainfall measurements on precipitation data files from the
\ac{NLDAS-2} re-analysis dataset, which is available for the continental
\ac{US}~\parencite{rui2013nldas}. We used data that were previously aggregated
from this re-analysis dataset to the county level by two of the coauthors
(Al-Hamdan and Crosson) for the Centers for Disease Control and Prevention's
\ac{WONDER} database \parencite{cdcwonder, alhamdan2014environmental}. To
create these aggregated data, these coauthors took hourly precipitation
measurements in the \ac{NLDAS-2} precipitation files, which are given
at~1/8\si{\degree} grid points across the continental \ac{US}, and summed them
by day for each grid point.  This created a daily precipitation total for each
grid point; these were then averaged for all grid points within a county's~1990
\ac{US} Census boundaries~\parencite{alhamdan2014environmental, cdcwonder}.
This process generated daily county-level precipitation estimates for each
continental \ac{US} county for 1988\,--\,2011 \parencite{cdcwonder}.  We
matched these county-level daily measurements by date with storm tracks, using
the date when the storm was closest to each county. Given the location of
storm-affected counties and the typical timing of tropical cyclones, these
precipitation measures primarily represent rainfall, although occasionally they
may represent snowfall or other types of precipitation.

@@

To validate these modeled county-level peak wind estimates, we compared them
to the  post-season wind radii values, which have been routinely published
since 2004 \parencite{knaff2016using}. These wind radii estimate the maximum
distance from the storm's center that winds of a certain intensity extend. They
give separate estimates for four quadrants of the storm to capture asymmetry in
the storm's wind field and are based on a post-season re-analysis that
incorporates all available data (e.g., satellite data, aircraft reconnaissance and
ground-based data if available) \parencite{knaff2016using}. Wind radii are
estimated for three thresholds of peak sustained surface wind: 64,~50, and~34
kt.  They therefore allow for the classification of counties into four
categories of peak sustained surface wind: $<$34~kt; 34\,--\,49.9~kt;
50\,--\,63.9~kt; and $\ge$64~kt.

We interpolated the wind radii data to~15-minute increments and
classified a county as exposed to winds in a given wind speed category if its
population mean center was within~85\% of the maximum radius for that wind
speed in that quadrant of the storm. The~85\% adjustment is based on previous
research that found that this ratio helps capture average wind extents within a
quadrant from these maximum wind radii \parencite{knaff2016using}.  We compared
these wind radii--based estimates of county-level peak sustained surface wind
during a storm with the modeled wind estimates, comparing all study storms
since 2004 for which at least one study county experienced peak sustained winds
of~$\ge$34 kt (based on the \ac{HURDAT2} wind radii). For each of these storms,
we calculated the percent of study counties that were classified in the same
wind category based on both data sources.

@@

To identify flood- and tornado-based tropical cyclone exposures in \ac{US}
counties, we matched storm tracks with event listings from the National Oceanic
and Atmospheric Administration (NOAA)'s Storm Event
Database~\parencite{stormevents}. While this database has recorded storm data,
particularly tornadoes, since~1950, its coverage changed substantially in~1996
to cover more types of storm events, including flood
events~\parencite{stormevents}. We therefore only considered flood metrics of
hurricane exposure for tropical cyclones in~1996 and later.

@@

The tornado observations from this dataset form a traditional tornado event
database for the \ac{US}, and so we did not further validate
the tornado event data. It is difficult to characterize flooding at the county
level because flooding can be very localized and can be triggered by a variety
of causes. To investigate the extent to which the NOAA flood event listings
capture extremes that might be identified with other flooding data sources, we
investigated a sample of study counties, comparing the flood event data during
tropical storms with streamflow measurements at \ac{US} Geological Survey
county streamflow gages \parencite{usgsgages, countyfloods, dataRetrieval}.  
We considered nine study counties, selecting counties geographically spread
through storm-prone areas of the eastern \ac{US} and with multiple streamflow
gages reporting data during events. The sample counties were: Baltimore County,
MD; Bergen County, NJ; Escambia County, FL; Fairfield County, CT; Fulton
County, GA; Harris County, TX; Mobile County, AL; Montgomery County, PA; and
Wake County, NC. For each county, we first identified all streamflow gages in
the county with complete data for Jan.~1~1996\,--\,Dec.~31,~2018. If a storm
did not come within~500~\si{\kilo\metre} of a county, it was excluded from this
analysis, but all other study storms were considered. 
For each storm and county, we summed streamflow measurements across all county
gages to generate daily totals for the five-day window around the storm's
closest approach. We took the maximum of these daily streamflow totals as a
measure of the county's maximum daily streamflow during that storm. We also
calculated the percent of streamflow gages in the county with a daily
streamflow that exceeded a threshold of flooding (the streamgage's median value
for annual peak flow) on any day during the five-day
window. We investigated how these measurements varied between storms with
associated flood events listings versus storms without an
event listing for the county, to explore if storms with flood event listings
tended to be associated with higher streamflow at gages within the county.

@@

Two of the exposure metrics (flood- and tornado-based) were inherently binary
in our data, since these metrics were based on whether an event was listed in
the NOAA Storm Events database.  For other exposure metrics, each county was
classified as exposed to a tropical cyclone based on whether the exposure
metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
picked reasonable thresholds (e.g., the threshold for gale-force wind for the
wind exposure), but others could be used with the open-source data and its
associated software.

@@

We characterized patterns in county-level exposure in the eastern \ac{US} for
exposure assessment based on each of the measured tropical cyclone hazards
(cumulative precipitation, peak sustained winds, flooding, and tornadoes).
Depending on available exposure data, this assessment included some or all of
the period from~1988 to~2018 (Table 1). For each binary metric of exposure to a
tropical cyclone hazard, we first summed the total number of county-level
exposures over available years and mapped patterns in
these exposures.

@@

We have posted study data as an open-source R package
\parencite{hurricaneexposuredata}. We have posted additional data and code at
\url{https://github.com/geanders/county_hurricane_exposure}.

>>>>

For our study domain, we used counties in the eastern half of the \ac{US}
(Figure~\ref{fig:hurrtracks}).  We included all tropical cyclones between 1988
and 2018 that were tracked in \ac{HURDAT2} \parencite{landsea2013} and came
within~250 \si{\kilo\metre} of at least one eastern \ac{US} county
(Figure~\ref{fig:hurrtracks}). We therefore including all land-falling or
near-land Atlantic basin tropical cyclones while excluding storms that never
neared the \ac{US}.  \subsection*{Distance-based exposure metric}

We first measured how close each storm's central track came to each county. We
used tracking data from \ac{HURDAT2}, which records the storm center's position
at the synoptic times of 6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC}.  We
interpolated this position to~15-\si{\minute} intervals using natural cubic
spline interpolation~\parencite{hurricaneexposure}. At each~15-\si{\minute}
interval, we measured the distance between the storm's center and each county's
population mean center~\parencite{countycenters}. We took
the minimum distance for each county as a measure of how close the storm came
to the county over its lifetime.  We also recorded the time when this occurred
for each county so we could later link observed data on precipitation,
flooding, and tornadoes.  To allow matching with data recorded in
local time (e.g., health data), we converted these times from \ac{UTC} to local
time~\parencite{countytimezones}.

@@

We based storm rainfall measurements on precipitation data files from the
\ac{NLDAS-2} re-analysis dataset, which is available for the continental
\ac{US}~\parencite{rui2013nldas}. We used data that were previously aggregated
from this re-analysis dataset to the county level by two of the coauthors
(Al-Hamdan and Crosson) for the Centers for Disease Control and Prevention's
\ac{WONDER} database \parencite{cdcwonder, alhamdan2014environmental}. To
create these aggregated data, these coauthors took hourly precipitation
measurements in the \ac{NLDAS-2} precipitation files, which are given
at~1/8\si{\degree} grid points across the continental \ac{US}, and summed them
by day for each grid point.  This created a daily precipitation total for each
grid point; these were then averaged for all grid points within a county's~1990
\ac{US} Census boundaries~\parencite{alhamdan2014environmental, cdcwonder}.
This process generated daily county-level precipitation estimates for each
continental \ac{US} county for 1988\,--\,2011 \parencite{cdcwonder}.  In this
study, we matched these county-level daily measurements by date with storm
tracks, using the date when the storm was closest to each county. Given the
location of storm-affected counties and the typical timing of tropical
cyclones, these precipitation measures primarily represent rainfall, although
occasionally they may represent snowfall or other types of precipitation.

@@

To validate these modeled county-level peak wind estimates, we compared them
to the  post-season wind radii, which have been routinely published
since 2004 \parencite{knaff2016using}. These estimate the maximum
distance from the storm's center that winds of a certain intensity extend. They
give separate estimates for four quadrants of the storm to capture asymmetry in
the storm's wind field and are based on a post-season re-analysis that
incorporates all available data (e.g., satellite data, aircraft reconnaissance, and
ground-based data if available) \parencite{knaff2016using}. Wind radii are
estimated for three thresholds of peak sustained surface wind: 64,~50, and~34
kt.  They therefore allow for the classification of counties into four
categories of peak sustained surface wind: $<$34~kt; 34\,--\,49.9~kt;
50\,--\,63.9~kt; and $\ge$64~kt.

We interpolated the wind radii data to~15-minute increments using linear
interpolation \parencite{hurricaneexposuredata} and classified a county as
exposed to winds in a given wind speed category if its population mean center
was within~85\% of the maximum radius for that wind speed in that quadrant of
the storm. The~85\% adjustment is based on previous research that found that
this ratio helps capture average wind extents within a quadrant based on these
maximum wind radii \parencite{knaff2016using}.  We compared these wind
radii--based estimates of county-level peak sustained surface wind during a
storm with the modeled wind estimates, comparing all study storms since 2004
for which at least one study county experienced peak sustained winds of~$\ge$34
kt (based on the post-season wind radii). For each of these storms, we
calculated the percent of study counties that were classified in the same wind
category based on both data sources.

@@

To identify flood- and tornado-based tropical cyclone exposures in \ac{US}
counties, we matched storm tracks with event listings from the National Oceanic
and Atmospheric Administration (NOAA)'s Storm Event
Database~\parencite{stormevents}. While this database has recorded storm data,
particularly tornadoes, since~1950, its coverage changed substantially in~1996
to cover more types of storm events, including flood
events~\parencite{stormevents}. We therefore only considered flood metrics of
tropical cyclone exposure for storms in~1996 and later.

@@

The tornado observations from this dataset form a traditional tornado event
database for the \ac{US}, and so we did not further validate
the tornado event data. It is difficult to characterize flooding at the county
level because flooding can be very localized and can be triggered by a variety
of causes. To investigate the extent to which the NOAA flood event listings
capture extremes that might be identified with other flooding data sources, we
investigated a sample of study counties, comparing the flood event data during
tropical storms with streamflow measurements at \ac{US} Geological Survey
county streamflow gages \parencite{usgsgages, countyfloods, dataRetrieval}.  
We considered nine study counties, selecting counties geographically spread
through storm-prone areas of the eastern \ac{US} and with at least one streamflow
gage reporting data during all events. The sample counties were: Baltimore County,
MD; Bergen County, NJ; Escambia County, FL; Fairfield County, CT; Fulton
County, GA; Harris County, TX; Mobile County, AL; Montgomery County, PA; and
Wake County, NC.

For each county, we first identified all streamflow gages in the county with
complete data for Jan.~1~1996\,--\,Dec.~31,~2018. If a storm did not come
within~500~\si{\kilo\metre} of a county, it was excluded from this analysis,
but all other study storms were considered.  For each storm and county, we
summed streamflow measurements across all county gages to generate daily totals
for the five-day window around the storm's closest approach. We took the
maximum of these daily streamflow totals as a measure of the county's maximum
daily streamflow during that storm. We also calculated the percent of
streamflow gages in the county with a daily streamflow that exceeded a
threshold of flooding (the streamgage's median value for annual peak flow) on
any day during the five-day window. We investigated how these measurements
varied between storms with associated flood events listings versus storms
without an event listing for the county, to explore if storms with flood event
listings tended to be associated with higher streamflow at gages within the
county.

@@

Two of the exposure metrics (flood- and tornado-based) were inherently binary
in our data, since these metrics were based on whether an event was listed in
the NOAA Storm Events database.  For other exposure metrics, each county was
classified as exposed to a tropical cyclone based on whether the exposure
metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
picked reasonable thresholds (e.g., the threshold for gale-force wind for 
wind exposure), but others could be used with the open-source data and its
associated software.

@@

We characterized patterns in county-level exposure in the eastern \ac{US} for
exposure assessment based on each of the measured tropical cyclone hazards
(cumulative precipitation, peak sustained wind, flooding, and tornadoes).
Depending on available exposure data, this assessment included some or all of
the period from~1988 to~2018 (Table 1). For each binary metric of exposure to a
tropical cyclone hazard, we first summed the total number of county-level
exposures over available years and mapped patterns in
these exposures.

@@

We have posted most study data as an open-source R package
\parencite{hurricaneexposuredata}. We have posted remaining data and code at
\url{https://github.com/geanders/county_hurricane_exposure}.
